### 2022-03-17 ho-techniques/ho-techniques.tex
more

We focus on monomorphic higher-order logic, without the axiom of infinity or the
axiom of at least two individuals \cite{bm-14-automation-ho}. However, the
techniques can easily be extended with polymorphism. Indeed, Zipperposition
already supports some of them polymorphically. Further, we use exactly the same
notation for this logic as introduced in Chapter \ref{ch:pre}. Since we are
working with extensions of superposition, we assume a clausal structure
(Sect.~\ref{sec:pre:clauses}). Like in the previous chapter, literals
of clauses can contain arbitrary higher-order terms, including formulas.  
At CASC, most theorem provers, including Zipperposition, are invoked using a
sequence of different configurations (possibly in parallel) until either the
time limit is reached or a proof is found. We call this sequence a \emph{portfolio}.

@@

To support such streams in the given clause procedure, we extend it to
represent the proof state not only by the active ($\mathcal{A}$) and passive ($\mathcal{P}$) clause
sets, but also by a priority queue $\mathcal{Q}$ containing the inference streams,
similar to the ``to do'' set $T$ present in the abstract Zipperposition loop of 
Waldmann et al.\  \cite[Sect.~4]{wtrb-20-sat-framework}.
Each stream is associated with a weight, and $\mathcal{Q}$ is sorted in order of
increasing weight\confrep{}{, a low weight corresponding to a high priority}.
When they introduced \lsup, Bentkamp et al.\
\cite{bbtvw-21-sup-lam} described an older version of this extension. Here
we present a newer version in more detail, including heuristics to postpone
unpromising streams. The pseudocode of the modified procedure is as follows:
\newcommand\ParamMode{\ensuremath{K_{\mathrm{fair}}}}
\newcommand\ParamMaxStreams{\ensuremath{K_{\mathrm{best}}}}
\newcommand\ParamRetry{\ensuremath{K_{\mathrm{retry}}}}
\newcommand{\assign}[2]{\State \ensuremath{\mathit{#1} \gets #2}}
\newcommand{\assignSameLine}[2]{\ensuremath{\mathit{#1} \gets #2}}
\algrenewcommand\algorithmicindent{1em}
\begin{algorithmic}[0]
    \Function{ExtractClause}{$Q$, $\mathit{stream}$}
      \assign{maybe\_clause}{\text{pop and compute the first element of } \mathit{stream} }
      \If{$\mathit{stream}$ is not empty}
        \State add $\mathit{stream}$ to $Q$ with an increased weight
      \EndIf
      \State \Return $\mathit{maybe\_clause}$
    \EndFunction

@@

  \newpage
  \Function{FairProbe}{$Q$, $\mathit{num\_oldest}$}
    \assign{collected\_clauses}{\emptyset}
    \assign{oldest\_streams}{\text{pop } \mathit{num\_oldest} \text{ oldest streams from } Q}
    \For{$\mathit{stream}$ in $\mathit{oldest\_streams}$}
      \assign{collected\_clauses}{\mathit{collected\_clauses} \mathrel\cup \textsc{ExtractClause}(Q, \mathit{stream})}
    \EndFor
    \State \Return $\mathit{collected\_clauses}$
  \EndFunction

>>>>

We focus on monomorphic higher-order logic, defined in Sect.~\ref{sec:pre:hol}. However, the
techniques can easily be extended with rank-1 polymorphism \cite{ksr-16-th1}. Indeed, Zipperposition
already supports some of them polymorphically. Further, we use exactly the same
notation for this logic and superposition calculus as introduced in Chapter \ref{ch:pre}. Since we are
working with extensions of superposition, we assume a clausal structure
(Sect.~\ref{sec:pre:clauses}). Like in the previous chapter, literals
of clauses can contain arbitrary higher-order terms, including formulas.  
At CASC, most theorem provers, including Zipperposition, are invoked using a
sequence of different configurations (possibly in parallel) until either the
time limit is reached or a proof is found. We call this sequence a \emph{portfolio}.

@@

\looseness=-1
To support such streams in the given clause procedure, we extend it to
represent the proof state not only by the active ($\mathcal{A}$) and passive ($\mathcal{P}$) clause
sets, but also by a priority queue $\mathcal{Q}$ containing the inference streams,
similar to the ``to do'' set $T$ present in the abstract Zipperposition loop of 
Waldmann et al.\  \cite[Sect.~4]{wtrb-20-sat-framework}.
Each stream is given a weight, and $\mathcal{Q}$ is sorted in order of
increasing weight\confrep{}{, a low weight corresponding to a high priority}.
When they introduced \lsup, Bentkamp et al.\
\cite{bbtvw-21-sup-lam} described an older version of this extension. Here
we present a newer version in more detail, including heuristics to postpone
unpromising streams. The pseudocode of the modified procedure is as follows:
\newcommand\ParamMode{\ensuremath{K_{\mathrm{fair}}}}
\newcommand\ParamMaxStreams{\ensuremath{K_{\mathrm{best}}}}
\newcommand\ParamRetry{\ensuremath{K_{\mathrm{retry}}}}
\newcommand{\assign}[2]{\State \ensuremath{\mathit{#1} \gets #2}}
\newcommand{\assignSameLine}[2]{\ensuremath{\mathit{#1} \gets #2}}
\algrenewcommand\algorithmicindent{1em}
\begin{algorithmic}[0]
    \Function{ExtractClause}{$Q$, $\mathit{stream}$}
      \assign{maybe\_clause}{\text{pop and compute the first element of } \mathit{stream} }
      \If{$\mathit{stream}$ is not empty}
        \State add $\mathit{stream}$ to $Q$ with an increased weight
      \EndIf
      \State \Return $\mathit{maybe\_clause}$
    \EndFunction

@@

  \pagebreak[2]
  \Function{FairProbe}{$Q$, $\mathit{num\_oldest}$}
    \assign{collected\_clauses}{\emptyset}
    \assign{oldest\_streams}{\text{pop } \mathit{num\_oldest} \text{ oldest streams from } Q}
    \For{$\mathit{stream}$ in $\mathit{oldest\_streams}$}
      \assign{collected\_clauses}{\mathit{collected\_clauses} \mathrel\cup \textsc{ExtractClause}(Q, \mathit{stream})}
    \EndFor
    \State \Return $\mathit{collected\_clauses}$
  \EndFunction

### 2022-03-17 pre/pre.tex
more

In this thesis we use classic simply typed monomorphic higher-order logic with
the choice operator (and some variations of this logic). Assuming a set of
base types $T$ and a set of symbols $\Sigma$, let us
define terms and types. Types are either base types $\tau \in T$, or function
types $\tau \rightarrow \upsilon$ where both $\tau$ and $\upsilon$ are types.
When we write $\tau_1 \rightarrow \cdots \rightarrow \tau_n \rightarrow \upsilon$
we assume $\upsilon$ to be a base type ($\upsilon \in T$), by convention.
Each symbol $\cst{f} \in \Sigma$ is assigned a type.

>>>>

In this thesis we use classic extensional simply typed monomorphic higher-order
logic with the choice operator, but without the description operator, the axiom of infinity,
and the axiom of at least two individuals \cite[Sect.~3.5]{bm-14-automation-ho}.
This logic corresponds to Henkin's extensional type theory. We additionally use
some variations of this logic, clearly stating the differences. Assuming a set of
base types $T$ and a set of symbols $\Sigma$, let us define terms and types.
Types are either base types $\tau \in T$, or function types $\tau \rightarrow
\upsilon$ where both $\tau$ and $\upsilon$ are types. When we write $\tau_1
\rightarrow \cdots \rightarrow \tau_n \rightarrow \upsilon$ we assume $\upsilon$
to be a base type ($\upsilon \in T$), by convention. Each symbol $\cst{f} \in
\Sigma$ is assigned a type.
